Crowd density estimation from Wi-Fi positioning data in the Amsterdam Arena

### 2016-06-21 PaperOnSmoothing/draft-31052016.tex
introduction started

We present the Amsterdam Arena project, which involves observing and managing crowd behavior, using the Amsterdam Arena stadium as a living laboratory. The main scientific question we explore is how to detect anomalous behavior in large crowds in real time.

\begin{itemize}

>>>>

***********************************From Sonja. nb: the text will be polished in the final phases, first I care only about putting all ideas on paper***************

Crowd disasters have taken many human lives. The Love Parade disaster (Duisburg, 2010), the Ellis Park Stadium disaster (Johannesburg, 2001), the PhilSports Stadium stampede (Manila, 2006) are just a few recent examples. One of the reasons that disasters happen is a "lack of overview of everybody" [cite paper on love parade EPJ DS], that is, a lack of macroscopic overview of the crowd. Critical crowd density (cite paper 8/m2, find it in e-mails)) is a contributing factor to crowd disasters; yet, it is still challenging to determine timely when it occurs, so that a disaster can be prevented before it happens.

In our work we are interested in estimating crowd density during concerts in indoor spaces. A lot of research on estimating crowd density has been done using video processing from security cameras (cite papers (1), (2), (3) from proposal?). However, this approach does not apply in our case, because 1)  it is difficult to obtaine macroscopic overview ; 2) the lighting conditions during concert hours are not sufficient for video-based crowd analysis; and 3) the error of video-based density estimation increases with the increase of the actual crowd density.

In our approach, we combat the three above mentioned issues with video analysis by exploiting the ubiquity of smart phones, as in ((4) (5), (6) from proposal?). Our living laboratory is the Amsterdam Arena(citatioon). First, we are estimating in real-time anonymously the positions of the visitors via the signal strengths of their phones to the  Wi-Fi routers, as presented in (cite Jan's slides), via a technique similar to trilateration. Second, we apply the law of the large numbers to obtain an estimation of the crowd density, also in real-time. With the first step, we are dealing with issues 1) and 2) mentioned above. With the second step, we are handling issue 3) -  that is, with our method the precision of estimation actually increases as the crowd density increases.

n.b. in related work section put all that use wi-fi tracking, for those that use GPS tracking only mention that this does not work in indoor spaces and that it requires participation. Not necessary to put work on video tracking in related work - it is already mentioned in intro. Mention also bluetooth tracking, but bluetooth is not so ubiquitous and our principles apply also if bluetooth was used (right?)

Finally, none of the mentioned work uses our approach of modeling the position of an individual as a probability distribution; our method is designed to attack the problem of having a dense crowd; we are not interested in precise estimations for freely moving crowds; rather, we are interested in obtaining precise estimation when the  concert crowds is dense and static due to this. blah blah (sonja: explain better)

************************************end from Sonja
\begin{itemize}
